---
title: "p8105_HW1_jh4054"
author: "Joy Hsu"
date: "9/19/2018"
output: github_document
---

This R markdown document is for Data Science P8105 - Homework 1
```{r chunk_tidyverse}
#Load tidyverse package
library(tidyverse)
```

#Problem 1

The first objective is to create a data frame comprised of 4 variables

To start, I created four vectors:
* A random sample of size 10 from a uniform[0, 5] distribution
* A logical vector indicating whether elements of the sample are greater than 2
* A (length-10) character vector
* A (length-10) factor vector
```{r}
#set seed to ensure reproducibility 
set.seed(5)

#create four variable types
vec_num = runif(10, 0, 5)
vec_log = vec_num > 2
vec_char = c("there", "is", "an", "alien", "hiding", "behind", "the", "black", "berry", "tree")
vec_fac = factor(c("alien", "platypus", "alien", "platypus", "alien", "platypus", "platypus", "platypus", "platypus", "platypus"))
```

Using the "tibble" function, I bounded the four variables in a dataframe. 
```{r}
#create data frame
prob1_df = tibble(vec_num, vec_log, vec_char, vec_fac)
```

The mean function will only work for some of the variables. The mean function works for the numeric variable since the inputs are already numeric. The mean function works for the logical vector since the outputs "TRUE" and "FALSE" are coded as numeric values 0 and 1, respectively. The character vector "vec_char" and factor vector "vec_fact" have categorical variables which cannot be averaged numerically.

```{r, message = F, warning = F}
#take the mean of numeric vector
mean(vec_num)

#take the mean of logical vector
mean(vec_log)

#take the mean of charactor vector
mean(vec_char)

#take the mean of factor vector
mean(vec_fac)
```

We can convert the character "vec_char" and factor variables "vec_factor" into numeric variables using the "as.numeric" function
```{r}
num_vec_log = as.numeric(vec_log)
num_vec_char = as.numeric(vec_char)
num_vec_fac = as.numeric(vec_fac)
```


In a second code chunk, convert your character variable from character to factor to numeric; similarly, convert your factor variable from factor to character to numeric. What happens?
```{r}
char_fac_num = as.numeric(as.factor(vec_char))
fac_char_num = as.numeric(as.character(vec_fac))
```

#Problem 2
```{r}
x = rnorm(1000)
```

