---
title: "p8105_HW1_jh4054"
author: "Joy Hsu"
date: "9/19/2018"
output: github_document
---

This R markdown document is for Data Science P8105 - Homework 1
```{r chunk_tidyverse}
#Load tidyverse package
library(tidyverse)
```

#Problem 1

The first objective is to create a data frame comprised of 4 variables

To start, I created four vectors:
* A random sample of size 10 from a uniform[0, 5] distribution
* A logical vector indicating whether elements of the sample are greater than 2
* A (length-10) character vector
* A (length-10) factor vector
```{r}
#set seed to ensure reproducibility 
set.seed(5)

#create four variable types
vec_num = runif(10, 0, 5)
vec_log = vec_num > 2
vec_char = c("there", "is", "an", "alien", "hiding", "behind", "the", "black", "berry", "tree")
vec_fac = factor(c("alien", "platypus", "alien", "platypus", "alien", "platypus", "platypus", "platypus", "platypus", "platypus"))
```

Using the `tibble` function, I bounded the four variables in a dataframe. 
```{r}
#create data frame
prob1_df = tibble(vec_num, vec_log, vec_char, vec_fac)
```

The mean function will only work for some variables. The mean function works for the numeric variable since the inputs are already numeric. The mean function also works for the logical vector since the outputs `TRUE` and `FALSE` of the logic operator are coded as numeric values 0 and 1, respectively. The character vector `vec_char` and factor vector `vec_fac` have categorical variables which cannot be averaged numerically. Taking the mean for character and factor variables will return NAs.

```{r, message = F, warning = F}
#take the mean of numeric vector
mean(vec_num)

#take the mean of logical vector
mean(vec_log)

#take the mean of charactor vector
mean(vec_char)

#take the mean of factor vector
mean(vec_fac)
```

We can convert the factor vector `vec_fac` and logical vector `vec_log` into numeric variables using the `as.numeric` function. The character vector conversion did not work and returned `NAs introduced by coercion`
```{r, message = F, warning = F}
#attempt to convert variables to numeric class 
num_vec_log = as.numeric(vec_log)
num_vec_char = as.numeric(vec_char)
num_vec_fac = as.numeric(vec_fac)
```

By converting the character variable from charactor to factor to numeric, we can successfully convert the character vector to a numeric vector. Conversion of the factor vector from factor to character to numeric failed.    
```{r, message = F, warning = F}
char_fac_num = as.numeric(as.factor(vec_char))
fac_char_num = as.numeric(as.character(vec_fac))
```

#Problem 2
First, we will create a data frame `r prob2_df` 
```{r}
x = rnorm(1000)
y = rnorm(1000)
logic_sum = x + y > 0
num_logic_sum = as.numeric(logic_sum)
fac_logic_sum = as.factor(logic_sum)

prob2_df = tibble(x, y, logic_sum, num_logic_sum, fac_logic_sum)
```



